---
title: "Blog1 - United States"
description:
  Entry 1 - The update of COVID-19 in the United States.
author:
  - name: ZIYAO WANG (Billy)
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    toc: true
    toc_float: true
    self_contained: false
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F)
```

```{r loadpackages, eval=TRUE}
library(tidyverse)
library(DT)
library(kableExtra)
library(lubridate)
library(plotly)
library(sparkline)
```

## Target Country - United Stated

 - Brief Introduction: It has been a long time since the rapid spread of coronavirus, it creates many inconveniences for people, more importantly, threatening our physical and psychological health. While projecting ourselves safe from the infection, it is also valuable to obtain updated information to inform ourselves of the latest progress since this is a fight for all mankind. Thus, this blog will introduce an overview of the COVID-19 situation in the USA with associated details.   

### Data Description

This blog has used two main data sets as the base for the construction of data communication regarding COVID-19 information.

#### Data Source 1 - R Package `nCov2019`

The data set is collected from various sources for different relevant information.  
 - Latest data of corona virus cases from [worldometers](https://www.worldometers.info/coronavirus/)  
 - Historical data of corona virus cases from [JHUCSSE](https://coronavirus.jhu.edu/map.html)  
 - Vaccine data from [raps.org](https://www.raps.org/news-and-articles/news-articles/2020/3/covid-19-vaccine-tracker)  
 - Therapeutics data from [raps.org](https://www.raps.org/news-and-articles/news-articles/2020/3/covid-19-therapeutics-tracker)  

This data source contains two core parts. The first one is the GUI explorer section which has a function of the `dashboard` to use a simple and interactive dashboard for information communication. 

The second part is the CLI analysis which includes a workflow from data query to data operation and lastly the visualization output of both geographic and animation plots.

This blog will __*focus on the CLI analysis part*__, which will query the updated COVID-19 data from five different aspects. It will include the "latest update COVID-19 info.", "global COVID-19 info.", "historical COVID-19 info.", "vaccine details", "therapeutics status and progress". 

<center>**Here are the variable descriptions for this data source**</center>

```{r variablenames, eval=TRUE}
variable <- data.frame(
  'variable_names' = c(
    "active number",
    "activePerOneMillion",
    "cases",
    "casesPerOneMillion",
    "continent",
    "country",
    "critical",
    "criticalPerOneMillion",
    "date",
    "deaths",
    "deathsPerOneMillion",
    "oneCasePerPeople",
    "oneDeathPerPeople",
    "oneTestPerPeople",
    "population",
    "recovered",
    "recoveredPerOneMillion",
    "tests",
    "testsPerOneMillion",
    "todayCases",
    "todayDeaths",
    "todayRecovered",
    "updated"
  ),
  'description' = c(
    "confirmed cases - deaths - recovered",
    "active number / million population",
    "confirmed cases",
    "confirmed cases / million population",
    "continent",
    "country",
    "Critical patients",
    "Critical patients / million population",
    "date",
    "deaths",
    "deaths patients / million population",
    "one Case Per People",
    "one Death Per People",
    "one Test Per People",
    "population",
    "recovered",
    "recovered Per One Million",
    "COVID-19 test",
    "COVID-19 test / million population",
    "confirmed cases in today",
    "confirmed death cases in today",
    "confirmed recovered cases in today",
    "the latest update time"
  )
)

```

```{r variable-table, eval=TRUE}
variable %>% 
  kable(caption = "Variable Summary for Data Source 1") %>% 
  kable_classic(full_width = F, html_font = "Cambria") %>% 
  kable_styling(bootstrap_options = c("striped","hover", "condensed"),
                position = "center",
                font_size = 10) %>% 
  row_spec(0, bold = T, color = "#4199C4", background = "#A7F5D0") %>% 
  row_spec(0:23, align = "c") %>% 
  column_spec(1:2, border_right = T, border_left = T) %>% 
  scroll_box(height = "200px")
```



#### Data Source 2 - [New York Times Covid-19 data sets](https://github.com/nytimes/covid-19-data)

This data source is collected by the journalists from the New York Times across several time zones to monitor news conferences and public data releases with seeking clarification from official bodies. It is under the license of [CC BY-NC](https://creativecommons.org/licenses/by-nc/4.0/legalcode) which allows the users to adapt the data with proper attribution given.  

> Contact Information: covid-data@nytimes.com

<center>**Here are the variable descriptions for this data source**</center>

```{r variablenames2, eval=TRUE}
variable2 <- data.frame(
  "variable_name" = c(
    "date",
    "state",
    "fips",
    "cases",
    "deaths"
  ),
  types = c(
    "character",
    "character",
    "integer",
    "integer",
    "integer"
  ),
  description = c(
    "date of record",
    "state name (in US)",
    "Federal Information Processing Standards (unique code for each state/counties)",
    "confirmed cases (accumulated)",
    "confirmed deaths (accumulated)"
  )
)

variable2 %>% 
  kable(caption = "Variable Summary for Data Source 2") %>% 
  kable_classic(full_width = F, html_font = "Cambria") %>% 
  kable_styling(bootstrap_options = c("striped","hover", "condensed"),
                full_width = F,
                position = "center",
                font_size = 10) %>% 
  row_spec(0, bold = T, color = "#4199C4", background = "#A7F5D0", font_size = 12) %>% 
  row_spec(0:5, align = "c") %>% 
  column_spec(1:3, border_right = T, border_left = T)
```


### Data Exploratory 

```{r loaddata, include=FALSE, message=FALSE, echo=FALSE}
# remotes::install_github("yulab-smu/nCov2019", dependencies = TRUE, force = TRUE)
library(nCov2019)
res <- query()
```

```{r readdata,message=FALSE, echo=FALSE}
x <- res$latest
y <- res$historical

```


```{r data1}
df_brief <- x["Global"]
df_detail <- (x$detail)
us_his <- y["USA"]

```

#### Exploratory 1 - Globel Overview

This table \@ref(fig:table1) has illustrated the updated information for top 50 countries _**by the total confirmed cases at 01/09/2021**_. It includes the daily confirmed cases, deaths and recovered (at 01/09/2021), etc. It is obvious that the USA is having the most confirmed cases at this date, hence it may indicate the epidemic circumstance is tough and complicated there.


```{r table1, fig.cap="table1"}
top50 <- df_brief %>% 
  arrange(desc(cases)) %>% 
  slice_head(n=50) %>% 
  select(-updated, -(cases:recovered))

table1 <- top50 %>% 
  arrange(desc(todayCases))

datatable(table1, 
           escape = FALSE,
           filter = 'top',
           caption = "COVID-19 updates for Top 50 countries (by total confirmed cases)",
            colnames = c("Country","Total Active cases",
                         "Update confirmed cases", "Update deaths", 
                         "Update recovered", "Total population", "Total tested"),
            options = list(pageLength = 10,
                           lengthMenu = c(10, 25, 50))) %>% 
  formatStyle('country', backgroundColor = 'steelblue', fontWeight = 'bold')

```

#### Exploratory 2 - US Focus


The table \@ref(fig:table2-sparkline) shows the monthly movement of USA's daily confirmed cases, death and recoveries. 

Something to notice:
 1. The number of deaths is decreasing over time, observing it through checking the numbers under `New Deaths`.   
 2. Based on the daily number of confirmed new cases, the pandemic was under controlled in June and August 2020, yet quickly rebounded back from October. In May 2021, things were under control again yet not for long, another breakout started in July. 
 
 - The possible cause for this repetitive fluctuation on the COVID-19 control in the USA may be caused by various reasons. One of them may be the [US President Election](https://time.com/tag/2020-election/), which made a change to the government body and led to different implementations against the pandemic issues.  
 
 - Another reason may be that many civil rights protests were held during the time, which may result in a lagged rapid increase in the confirmed cases after the event.  [Examples - BLM](https://www.nytimes.com/interactive/2020/07/03/us/george-floyd-protests-crowd-size.html) 


 * _There is a stop of daily recoveries report at 13/12/2020_, it may cause by either stop record of the data source, or miss recording.  

<center>**Monthly COVID-19 updates for USA**</center>


```{r table2-sparkline, fig.cap="table2"}
table2 <- us_his %>% 
  mutate(year = year(date),
         month = month(date),
        new_cases = cases - lag(cases),
        new_deaths = deaths - lag(deaths),
        daily_recovered = recovered - lag(recovered)
        ) %>% 
  group_by(year,month) %>% 
  summarise(new_cases = spk_chr(new_cases,
                                type = "line"),
            new_deaths = spk_chr(new_deaths,
                                 type = "line"),
            daily_recovered = spk_chr(daily_recovered,
                                 type = "line")
            ) 

 datatable(table2, 
           escape = FALSE,
           filter = 'top',
           caption = "Monthly variations in confirmed, deaths, and recovered cases in USA",
            colnames = c("Year","Month", "New Cases", "New Deaths", "Daily recoveries"),
            options = list(lengthMenu = c(10, 25),
              fnDrawCallback = 
            htmlwidgets::JS('function(){HTMLWidgets.staticRender();}'))) %>% 
    spk_add_deps()
```

#### Exploratory 3 - US's State-level

This table \@ref(fig:table3) demonstrates the updated COVID-19 information on `01/09/2021` for each state (included some special areas, such as Guam and Puerto Rico). It shows a trend that the confirmed cases are centralized in the big cities and states, which may indicate the need for additional supports for these states to help to control the spread of coronavirus over the country effectively.


<center>**2021/09 monthly COVID-19 updates for each states in USA**</center>


```{r data2}
state_data <- read.csv(here::here("data/us-states.csv"))
update_state <- state_data %>% 
  mutate(year = year(date),
         month = month(date),
         states = state,
         'cases(accumulated)' = cases,
         'deaths(accumulated)' = deaths) %>% 
  select(year:`deaths(accumulated)`) %>% 
  filter(year == "2021") %>% 
  filter(month == 9)

```

```{r table3, fig.cap="table3"}
table3 <- update_state %>% 
  arrange(desc(`cases(accumulated)`)) %>% 
  select(-year, -month) %>% 
  datatable(filter = 'top', 
           escape = FALSE,
           caption = "Monthly variations in confirmed, deaths, and recovered cases in USA",
            colnames = c("states", "Total confirmed cases", "Total deaths"),
            options = list(pageLength = 10,
                           lengthMenu = c(10, 25, 35,55))
           )%>% 
  formatStyle('states', fontWeight = 'bold')
table3

```

This interactive figure \@ref(fig:figure1) shows the top 5 states with the greatest total confirmed cases movement from 2020 to 2021. It may help to understand the trend mentioned in the last part better with details.  

```{r figure1, preview = TRUE, fig.cap="Movement of top 5 states' total confirmed cases"}
top5_state <- state_data %>% 
  filter(state %in% c("California",
                      "Texas",
                      "Florida",
                      "New York",
                      "Illinois"
                      )) %>% 
  select(-fips) %>% 
  mutate(date = ymd(date))


figure1 <- top5_state %>% 
  arrange(desc(cases)) %>% 
  ggplot(aes(x = date, y = cases, color = state))+
  geom_line()+
  theme_bw()
ggplotly()
```

**If you like this blog, subscribing it now! **

