---
title: "ETC5523: Communicating with Data"
subtitle: "Fashionable Model Output and Data Tables - Part B"
author: "Emi Tanaka"
email: "ETC5523.Clayton-x@monash.edu"
date: "Week 3"
length: "50 minutes"
department: "Department of Econometrics and Business Statistics"
titlebgimg: "images/bg-04.png"
output:
  xaringan::moon_reader:
    css:
      - ninjutsu 
      - "assets/font-awesome-all.css"
      - "assets/tachyons-addon.css"
      - "assets/animate.css"
      - "assets/fira-code.css"
      - "assets/boxes.css"
      - "assets/table.css"
      - "assets/styles.css"
      - "assets/monash-brand.css"
      - "assets/monash-fonts.css"
      - "assets/slide-types.css"
      - "assets/lists.css"
      - "assets/custom.css"
    self_contained: false 
    seal: false 
    chakra: 'lib/remark-latest.min.js'
    includes:
      in_header: "assets/head.html"
    lib_dir: lib
    #includes:
    #  in_header: "assets/custom.html"
    mathjax: "lib/mathjax-local/MathJax.js?config=TeX-AMS-MML_HTMLorMML"
    nature:
      highlightStyle: magula
      highlightLanguage: r 
      highlightLines: true
      highlightSpans: false 
      countIncrementalSlides: false
      slideNumberFormat: '%current%/%total%'
      navigation:
        scroll: false 
        touch: true
        click: false
      ratio: '16:9'
---

```{r, include = FALSE}
library(tidyverse)
library(janitor)
library(kableExtra)

current_file <- knitr::current_input()
basename <- gsub(".Rmd$", "", current_file)

knitr::opts_chunk$set(
  fig.path = sprintf("images/%s/", basename),
  fig.width = 6,
  fig.height = 4,
  fig.align = "center",
  fig.retina = 3,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  cache = FALSE,
  cache.path = "cache/"
)
```

```{r titleslide, child="assets/titleslide.Rmd"}
```

---

# <i class="fas fa-table"></i> Some common tables 

.grid[
.item.br[
.f3[Descriptive summary statistics tables [<i class="fas fa-link"></i>](http://thatdatatho.com/2018/08/20/easily-create-descriptive-summary-statistic-tables-r-studio/)]

<center style="padding:10px">
```{r, echo = FALSE}
df <- mtcars %>% mutate(cyl = factor(cyl),
                        vs = factor(vs))
table1::table1(~ mpg + wt + vs | cyl, data = df)
```
</center>

]
.item.pl3[
.f3[Regression results tables]
.f5[
```{r}
df <- mutate(mtcars, 
             cyl = as.factor(cyl))
fit1 <- lm(mpg ~ wt + vs + cyl, data = df)
fit2 <- lm(mpg ~ wt + cyl,  data = df)
```
]

<br>

```{r, echo = FALSE, results = "asis"}
t1 <- gtsummary::tbl_regression(fit1)
t2 <- gtsummary::tbl_regression(fit2)
gtsummary::tbl_merge(tbls = list(t1, t2),
                     tab_spanner = c("`fit1`", "`fit2`"))
```
]
]

---

class: middle

#  What statistics to present? 

--

<br>
.info-box[
.monash-blue[**This depends on what you want to convey and your audience.**]

There are two key purposes of the table:

1. _display_ information; and
2. _communicate_ information.


]

---

class: transition

# Types of tables

---

# <i class="fas fa-table red"></i> Descriptive summary statistics .f4[Part 1/2]

* The main goal is to give a numerical summary to give a "feel" of what the data contains. These generally should convey:
  * variables in the data,
  * the number of observations for each variable, 
  * missing values (if any),
  * the distribution, e.g. in the form of five number of summaries or counts/percentages for categorical variables. 
--

* For numerical variables, you may have a **correlation table** which displays the correlation coefficients of every pair of variables. Because it is symmetrical, you can omit out the upper triangle.

```{r, echo = FALSE}
cc <- cor(select(mtcars, mpg, wt, hp))
cc[upper.tri(cc, diag = T)] <- NA
tt <- as_tibble(cc)
tt %>% mutate(Variable = c("1. Miles per gallon", "2. Weight", "3. Horsepower"),
              Mean = round(c(mean(mtcars$mpg), mean(mtcars$wt), mean(mtcars$hp)), 2),
              SD = round(c(sd(mtcars$mpg), sd(mtcars$wt), sd(mtcars$hp)), 2),
              mpg = ifelse(is.na(mpg), "", round(mpg, 2)),
              wt = ifelse(is.na(wt), "", round(wt, 2))) %>% 
  select(Variable, Mean, SD, mpg, wt) %>% 
  knitr::kable(table.attr = "class='cor-table'",
               col.names = c("Variables", "Mean", "SD",
                             "1.", "2.")) %>% 
  kable_classic(full_width = FALSE)
```



---

# <i class="fas fa-table red"></i> Descriptive summary statistics .f4[Part 2/2]

* You may have **cross-tabulations** (also called **contingency tables**)
```{r xtab, echo = FALSE}
mtcars %>% 
  tabyl(cyl, gear) %>% 
  adorn_totals(where = c("row", "col")) %>% 
  adorn_percentages("all") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns() %>% 
  adorn_title("combined") %>% 
  knitr::kable(table.attr = "class='xtab'") %>% 
  kable_styling() %>% 
  row_spec(4, extra_css = "border-top: 1px solid black;") %>% 
  kable_classic(full_width = FALSE)
```

---

# <i class="fas fa-table red"></i> Regression results 

.w-70[
* Or more generally, the important numerical characteristics of your model. This may include:
  * parameter estimates of your model,
  * the standard error or confidence/credible interval of your estimates,
  * model fit quality, e.g. \\(R^2\\), AIC, BIC and so on.
]
--

.w-70[
* You see often the inclusion of $p$-value, usually from the significance test of the corresponding variable and some indicate the significance level by the amount of `*`.
]
--

.w-70[
* It is important to **_convey the uncertainty_** for the estimates or predictions from your model. 
]
--

.w-70[
* There is also what is called the **ANOVA table** that shows the variation according to different sources .f4[(out of scope for this unit)].

]


---

class: transition 

# Communication <Br>considerations<br> for tables

---


# <i class="fas fa-syringe" style="color:#FFC0CB"></i> Numerical precision .f4[Part 1/3]

* Select an appropriate precision for your goal and audience:

```{r, echo = FALSE}
set.seed(1)
mtcars %>% 
  rownames_to_column() %>% 
  # add some random decimal places
  mutate(wt1 = round(wt + rnorm(n())/ 1000, 5),
         wt2 = round(wt, 2),
         wt3 = round(wt, 0)) %>%  
  select(rowname, wt1:wt3) %>% 
  head(4) %>% 
  knitr::kable(table.attr="class='xtab'",
        col.names = c("Car brand", "5 d.p.", "2 d.p.", "0 d.p.")) %>% 
  kable_styling() %>% 
  column_spec(2:4, width = "100px") %>% 
  add_header_above(c(" " = 1, "Weight" = 3)) %>% 
  kable_classic(full_width = FALSE)
```

---

# <i class="fas fa-syringe" style="color:#FFC0CB"></i> Numerical precision .f4[Part 2/3]

* Display trailing zeroes to match selected precision of the column:

```{r, echo = FALSE}
tibble(zero = c(0.233, 0.32, 0.4, 0.343)) %>% 
  mutate(nozero = as.character(zero)) %>% 
  knitr::kable(table.attr="class='xtab'", 
               align = "rr",
               col.names = c("Yes", "No")) %>% 
  kable_styling() %>% 
  column_spec(1:2, width = "100px") %>% 
  add_header_above(c("Trailing zeroes" = 2),
                   extra_css = "border-bottom: 1px solid black;") %>% 
  kable_classic(full_width = FALSE)
```

---

# <i class="fas fa-syringe" style="color:#FFC0CB"></i> Numerical precision .f4[Part 3/3]

* Change and display units as appropriate:

```{r, echo = FALSE}
mtcars %>% 
  rownames_to_column() %>% 
  # add some random decimal places
  mutate(wt1 = wt * 1000,
         wt2 = wt,
         wt3 = round(wt1 * 0.453592),
         wt4 = signif(wt3 * 1000, 3),
         wt5 = scales::comma(signif(wt3 * 1000000))) %>%  
  select(rowname, wt1:wt5) %>% 
  head(4) %>% 
  knitr::kable(table.attr="class='xtab'",
               col.names = c("Car brand", "Weight (lbs)", "Weight (1000 lbs)", "Weight (kg)", "Weight (g)", "Weight (mg)"),
               align = "lrrrrr") %>% 
  kable_styling() %>% 
  column_spec(2:6, width = "200px") %>% 
  kable_classic()
  
```
--

<Br>
* Show comma every 3 digits (or other marks as needed). <br>E.g. `1000000` is harder to read than `1,000,000`.

```{r}
scales::comma(c(439024, 4900343), accuracy = 1000)
```

---

# <i class="fas fa-align-left"></i> Column <i class="fas fa-align-center"></i> alignment   <i class="fas fa-align-right"></i>

* Spanner labels are usually aligned in center.
* Right-align numbers 
* Left-align texts

<Br>

```{r, echo = FALSE}
mtcars %>% 
  rownames_to_column("car") %>% 
  # add some random decimal places
  mutate(
    car1 = car,
    car2 = car,
    car3 = car,
    hp1 = hp,
    hp2 = hp,
    hp3 = hp) %>%  
  select(car1:hp3) %>% 
  head(4) %>% 
  knitr::kable(table.attr="class='xtab'",
               col.names = c("Left", "Center", "Right", "Left", "Center", "Right"),
               align = "lcrlcr") %>% 
  kable_styling() %>% 
  column_spec(1:3, width = "200px") %>% 
  column_spec(4:6, width = "80px") %>% 
  add_header_above(c("Car brand" = 3,
                     "Horsepower" = 3)) %>% 
  kable_classic()
```

---

# <i class="fas fa-align-justify"></i> Labels within tables

* It is possibly obvious, but tables designed as final produce (e.g. in report) should have polished labels
* For columns, the unit may be written in the column header label
* You shouldn't label the unit within the table

.flex[
.w-50[

.center.f1[
✅
]

```{r, echo = FALSE}
mtcars %>% 
  rownames_to_column("car") %>% 
  select(car, wt, disp) %>% 
  head() %>% 
  knitr::kable(col.names = c("Car brand", "Weight<br> (1000 lbs)", "Displacement<br> (cubic inches)"), escape = FALSE) %>% 
  kable_classic(full_width = FALSE)
```

]
.w-50[

.center.f1[
❌
]

```{r, echo = FALSE}
mtcars %>% 
  rownames_to_column("car") %>% 
  select(car, wt, disp) %>% 
  mutate(wt = paste(wt*1000, "lbs"),
         disp = paste(disp, "cubic inches")) %>% 
  head() %>% 
  knitr::kable() %>% 
  kable_classic(full_width = FALSE)
```
]
]


---

# <i class="fas fa-align-justify"></i> Texts accompanying tables

.w-70[
* Besides the contents of table, a table may be accompanied with: **_table header_**, **_caption_**, **_footnotes_** and/or **_source notes_**.
* The conventions of how and what to write will depend on your audience and medium of report
* Generally if you are *communicating* information, your caption should:
  * .monash-blue[summarise the take-away message], in other words, why should the audience care about this table?
  * give context of the table (e.g. " $R_0 > 1$ means that the virus is more infectious")
]


---

class: transition 


# Making tables with R 


---

# 🏗️ How to make tables in R?

.w-70[
* There are many packages that make table in R, including ones that wrangle the data for you to make specialised table output. E.g. `knitr::kable`, `kableExtra`, `formattable`, `gt`, `DT`, `pander`, `xtable`, `stargazer`.
]
--

.w-70[
* You can read the documentation for each packages to make the table you want (I mainly use `knitr::kable`, `kableExtra` and `DT`).
]
--

.w-70[
* Whatever package you use, it's important that you understand the output and how it works with the medium you are trying to display the table.
]




---


# <i class="fas fa-file"></i> Table in Markdown

In markdown file:

<br>

```
First Header  | Second Header
------------- | -------------
Content Cell  | Content Cell
Content Cell  | Content Cell
```

<br>
*Possible* display:

<br>

First Header  | Second Header
------------- | -------------
Content Cell  | Content Cell
Content Cell  | Content Cell

---

# <i class="fas fa-file"></i> Table in HTML & PDF
.grid[.item.br[
HTML &rarr; Web browser
```html
<table>
<thead>
<tr>
<th>First Header</th> <th>Second Header</th>
</tr>
</thead>
<tbody>
<tr>
<td>Content Cell</td> <td>Content Cell</td>
</tr>
<tr>
<td>Content Cell</td> <td>Content Cell</td>
</tr>
</tbody>
</table>
```

]
.item[
LaTeX &rarr; PDF
```latex
\begin{tabular}{cc}
\hline
First Header & Second Header\\
\hline
Content Cell & Content Cell\\
Content Cell & Content Cell\\
\hline
\end{tabular}
```

{{content}}

]
]

--

* HTML is for HTML output 
* LaTeX is for PDF output
* Most HTML elements do _not_ work in LaTeX and vice versa.

---

layout: true

# Case study .circle.bg-black.white[2]  building static tables by components with `gt`

.flex[
.w-50[
<img src="images/lecture-03/gt_parts_of_a_table.svg">

.f4[Source: https://gt.rstudio.com/]
]
.w-50.f5[
{{content}}
]
]

---

---

```{r}
library(gt)
df <- select(head(mtcars), wt, disp, cyl)
```



---

```{r}
library(gt)
df <- select(head(mtcars), wt, disp, cyl)
gt(df) #<<
```

---

```{r}
library(gt)
df <- select(head(mtcars), wt, disp, cyl)
gt(df) %>% 
  tab_header(title = "Motor Trend Car Road Tests", #<<
             subtitle = "Design and performance of 1973-74 automobile models") #<<
```

---

```{r}
library(gt)
df <- select(head(mtcars), wt, disp, cyl)
gt(df) %>% 
  tab_header(title = "Motor Trend Car Road Tests",
             subtitle = "Design and performance of 1973-74 automobile models") %>% 
  tab_source_note(md("Source: 1974 *Motor Trend* US magazine")) #<<
```

---

```{r}
library(gt)
df <- select(head(mtcars), wt, disp, cyl)
gt(df) %>% 
  tab_header(title = "Motor Trend Car Road Tests",
             subtitle = "Design and performance of 1973-74 automobile models") %>% 
  tab_source_note(md("Source: 1974 *Motor Trend* US magazine")) %>% 
  cols_label( #<<
    wt = html("Weight<br>(1000lbs)"), #<<
    disp = html("Displacement<br> (inch<sup>3</sup>)"), #<<
    cyl = html("Number of<br>cylinders") #<<
  ) #<<
```


---

layout: false 

# Case study .circle.bg-black.white[3] interactive tables with `DT`

```{r}
DT::datatable(mtcars, options = list(pageLength = 4))
```

--

<div class="bg-white border-box" style="position:absolute;right:50px;bottom:50px;width:700px;padding:10px;">
<ul>
<li><code>DT::datatable</code> works best for HTML documents.</li> 
<li>For PDF documents, it takes a <code>webshot</code> of the HTML table and inserts it as an image.</li>
<li>It's useful for data exploration where the main goal of the table is display of information rather than communication.</li>
</ul>
</div>

---

class: middle

# When do you make tables over plots?

When you want to **_show exact values_** or the **_accuracy of the values are important_** to convey.

--

<br>

# You _can_ combine plots with tables.

---

# <i class="fas fa-table red"></i> Case study .circle.bg-black.white[4]  heatmap table with `formattable` 

```{r}
library(formattable)
mtcars %>% 
  select(mpg, disp, wt, hp) %>% 
  cor() %>% 
  as.data.frame() %>% 
  rownames_to_column("Variables") %>% 
  formattable(list(area(col = 2:5) ~ color_tile("#F5B7B1", "#7DCEA0")))
```


---

# <i class="fas fa-table red"></i> Case study .circle.bg-black.white[5]  inline plots with `sparkline` 


* **Sparkline** refers to a small chart drawn without axes or coordinates. 
* There is a jQuery sparkline that is provided as a `htmlwidgets` in the [`sparkline` R-package](https://github.com/htmlwidgets/sparkline).
--

* This is a `htmlwidgets` object so it only works for HTML output and requires the JS dependencies to work.

--


<center>
<a href="lecture-03suppB.Rmd" download><i class="fas fa-file"></i></a>
.flash.animated.blue[LIVE DEMO]
</center>
---

# <i class="fas fa-chalkboard-teacher bg-white animated bounceIn" style="padding:3px;border:1px dashed #0092EB; color: #0092EB;"></i> Today's main lessons 

# Working with model objects 

* Understanding about the OOP system
* How to tease apart the model objects and methods associated with it

# Presenting information as a table

* Various common tables that present information
* Some guidelines for best way to communicate with tables


---

```{r endslide, child="assets/endslide.Rmd"}
```
